---
title: "Google Analytics Customer Revenue Prediction"
output: html_notebook
---

# Introdcution

A detailed analysis for Google Merchandise Store (also known as GStore, where Google swag is sold) customer dataset to predict revenue per customer. The goal is to enable marketing teams who analyze data on top of Google Analytics data to make more actionable operational changes and investment decisions.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*. 

# Loading and Exploring Data

```{r}
# loading data from CSV files.
train <- read_csv("./data/input/train.csv")
test <- read_csv("./data/input/test.csv")

cat("Train dataset overview: \n\n")
glimpse(train)
cat("\n\nTest dataset overview: \n\n")
glimpse(test)
```

There are 12 features:

* **fullVisitorId** - an unique identifier for each user of the Google Merchandise Store
* **channelGrouping** - the channel via which the user came to the Store
* **date** - the date on which the user visited the Store
* **device** - the specifications for the device used to access the Store
* **geoNetwork** - this section contains information about the geography of the user
* **sessionId** - an unique identifier for this visit to the store
* **socialEngagementType** - engagement type, either "Socially Engaged" or "Not Socially Engaged"
* **totals** - this section contains aggregate values across the session
* **trafficSource** - this section contains information about the Traffic Source from which the session originated
* **visitId** - an identifier for this session
* **visitNumber** - the session number for this user
* **visitStartTime** - the timestamp (POSIX).

The columns device, geoNetwork, trafficSource, and totals are in JSON format. 
```{r}
library(jsonlite)

flatten_json <- . %>%
  str_c(., collapse = ",") %>%
  str_c("[", .,"]") %>%
  fromJSON(flatten = T)

parse <- . %>%
  bind_cols(flatten_json(.$device)) %>%
  bind_cols(flatten_json(.$geoNetwork)) %>%
  bind_cols(flatten_json(.$trafficSource)) %>%
  bind_cols(flatten_json(.$totals)) %>%
  select(-device, -geoNetwork, -trafficSource, -totals)

train <- parse(train)
test <- parse(test)

```

## Removing the identification variables

We have two identification variables; sessionId, visitId, and fullVisitorId. We will keep the fullVisitorId as it is needed in our end result and will remove the others from training and testing data.

```{r}
train <- train %>% select(-sessionId, -visitId)
test <- test %>% select(-sessionId, -visitId)
```


## Finding Constant Variables

```{r}

train %>%
  map_dfr(n_distinct) %>%
  gather() %>%
  ggplot(aes(reorder(key, -value), value)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  scale_y_log10(breaks = c(5, 50, 250, 500, 1000, 10000, 50000)) +
  geom_text(aes(label = value), vjust = 1.6, color = "white", size=3.5) +
  theme_minimal() +
  labs(x = "features", y = "Number of unique values") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).
